MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-06-26 latex/paper/sections/10-placement.tex
minor edits

Shown in Listing \ref{lst:sa_outer} is a simplified pseudocode for the outer-most loop of our placer's SA algorithm. 
First, we randomly place all \texttt{SiteInst}s across the \texttt{device}.
Then, we initialize a simple geometric cooling schedule: 
\begin{equation}
    \label{coolingSchedule}
    T_{n+1} = \alpha T_n
\end{equation}
with initial temperature \(T_0=10000\), cooling rate \(\alpha=0.98\), and \(n = \{0, 1, 2, ..., 300\}\) passes.
Hill-climbing moves can be accepted via the probability function \(P()\): 
\begin{equation}
    \label{moveAcceptance}
    P(s, s^*, T_n) = exp(\frac{E(s) - E(s^*)}{T_n})
\end{equation}
where \(E(s^*)\) is the energy of the proposed state \(s^*\), \(E(s)\) is the energy of the current state \(s\), and \(T_n\) is the current global temperature at \(n\) passes. 
If \(P(s, s^*, T_n) < random(0, 1)\), then \(s \leftarrow s^*\).
Recall that this probability is only evaluated for hill-climbing moves, thus the cost difference \(E(s) - E(s^*)\) is always negative.
In summary, a lower the cost difference and a higher the temperature contributes to a higher the probabilty of hill-climbing acceptance.
All hill-descending moves on the other hand are simply accepted outright.

@@

In force directed moves, there is a concept of global placement versus detailed placement, where global placement is driven by forces and detailed placement attempts to legalize the global placement by discretizing it onto the grid of \texttt{Site}s on the device.
There also exists other strategies to mix packing, placement, and legalization in different ways. (CITE THAT ONE PAPER).
In this context, our placer follows a straightforward pack-legalize-place strategy.

>>>>

Shown in Listing \ref{lst:sa_outer} is a simplified pseudocode for the outer-most loop of our placer's SA algorithm. 
First, we randomly place all \texttt{SiteInst}s across the \texttt{device}.
Then, we initialize a simple geometric cooling schedule: 
\begin{equation}
    \label{coolingSchedule}
    T_{n+1} = \alpha T_n
\end{equation}
with initial temperature \(T_0=10000\), cooling rate \(\alpha=0.98\), and \(n = \{0, 1, 2, ..., 300\}\) passes.
Hill-climbing moves can be accepted via the probability function \(P()\): 
\begin{equation}
    \label{moveAcceptance}
    P(s, s^*, T_n) = exp(\frac{E(s) - E(s^*)}{T_n})
\end{equation}
where \(E(s^*)\) is the energy of the proposed state \(s^*\), \(E(s)\) is the energy of the current state \(s\), and \(T_n\) is the current global temperature at \(n\) passes. 
If \(P(s, s^*, T_n) < random(0, 1)\), then \(s \leftarrow s^*\).
Recall that this probability is only evaluated for hill-climbing moves, thus the cost difference \(E(s) - E(s^*)\) is always negative.
In summary, a lower cost difference and a higher temperature contributes to a higher probabilty of hill-climbing acceptance.
All hill-descending moves on the other hand are simply accepted outright.

@@

In force directed moves, there is a concept of global placement and detailed placement, where global placement is driven by forces and detailed placement attempts to legalize the global placement by discretizing it onto the grid of \texttt{Site}s on the device.
In such placers, objects are first placed freely on an analog field in global placement, even allowing objects to overlap with one another. 
Then, in detailed placement, the object positions are discretized amongst the grid of resources and overlapping objects spread out until there are no overlaps.
In literature, detailed placement is also referred to as legalization.

There also exists other strategies to mix packing, placement, and legalization in different ways. (CITE THAT ONE PAPER).
In this context, our placer follows a straightforward pack-legalize-place strategy.

### 2025-06-26 latex/paper/sections/11-placement_results.tex
minor edits

Shown in figure \ref{fig:placers_overlay} are the total HPWL curves over number of passes for each of the five placers. 
\texttt{PlacerGreedyMidpoint} (in red) appears to perform the worst by a considerable margin with a final HPWL cost of 590K.
This is likely due to the lack of randomness in moves resulting in a lack of diversity in movement proposals combined with the greediness of the movement acceptance that prevents the system from escaping local minima.
This is supported by its HPWL curve (in red) showing a sharp drop in cost that quickly asymptotes just below the 600K mark, suggesting that the system crystallized very quickly into a local minimum and was unable to climb out of it.

Next is \texttt{PlacerAnnealMidpoint} with a final cost of 466K.
Like the previous placer, \texttt{PlacerAnnealMidpoint} lacks the diversity of random moves which contributes to faster crystallization into local minima.
However, this placer likely performed better than the previous as there is still some randomness in the move acceptance evaluation which gives way to occasional hill-climbing. 
Keep in mind that centroid moves do not necessarily decrease system cost, as the centroid move will decrease the HPWL cost of the nets on connected to the current \texttt{SiteInst}, but can potentially increase the cost of another group of nets should a swap be evaluated, which can potentially result in a total increase in system cost.

@@

>>>>

Shown in figure \ref{fig:placers_overlay} are the total HPWL curves over number of passes for each of the five placers. 
\texttt{PlacerGreedyMidpoint} (in red) appears to perform the worst by a considerable margin with a final HPWL cost of 590K.
This is likely due to the lack of randomness or lack of diversity in move proposals combined with the persistent greediness of the movement acceptance that prevents the system from escaping local minima.
This conclusion is supported by its HPWL curve (in red) showing a sharp drop in cost that quickly asymptotes just below the 600K mark, suggesting that the system crystallized very quickly into a local minimum and was unable to climb out of it.

Next is \texttt{PlacerAnnealMidpoint} with a final cost of 466K.
Like the previous placer, \texttt{PlacerAnnealMidpoint} lacks the diversity of random moves which contributes to faster crystallization into local minima.
However, this placer likely performed better than the previous as there is still some randomness in the move acceptance evaluation which gives way to occasional hill-climbing. 
Keep in mind that centroid moves do not necessarily decrease system cost. 
The centroid move can decrease the HPWL cost of the nets on connected to the current \texttt{SiteInst}, but can potentially increase the cost of another group of nets in the event of a swap, which can potentially result in an increase in total system cost.

@@

In between \texttt{AnnealRandom} and \texttt{AnnealHybrid} comes \texttt{PlacerGreedyRandom} with a final cost of 333K.
